Enhancing instance-level constrained clustering through differential evolution

### 2019-06-11 Papers/DILS/paperDILS.tex
SHADE terminado y DILS avanzado

One of the most widely known and studied data analysis problems is clustering. It is one of the most successful methods in the field of unsupervised learning, where there is no supervision on how the information should be handled. Clustering has been able to provide solutions in a large number of knowledge fields, such as, marketing, banking, psychology, psychiatry, astronomy, archaeology, genetics, cartographic labeling and image segmentation among others \cite{Everitt:2009:CA:1538772, araujo2019improving, verma2016improved, aparajeeta2016modified, wang2018non}.

In the literature, clustering methods are often divided into two subsets: partitional clustering and hierarchical clustering. The main difference is that in hierarchical clustering the result is not a partition of the data with a certain number of clusters---as in partitional clustering---, but instead a dendrogram in which there are partitions that include from the whole dataset to particular individuals \cite{Everitt:2009:CA:1538772}. A representative example of partitional clustering is the widely studied and well-known algorithm K-means, while for hierarchical clustering the CURE method should be highlighted \cite{wu2009top, guha1998cure}. In this paper we will focus on partitional clustering.

We can define partitional clustering as the task of grouping the instances of a dataset into $k$ clusters, so that new information can be extracted from them. A dataset $X$ is composed of $n$ instances, each one of them described by $d$ features. More formally, $X = \{x_1, \cdots, x_n\}$, with the $i$th instance noted as $x_i = (x_{[i,1]}, \cdots, x_{[i,d]})$. A typical clustering algorithm assigns a class label $l_i$ to each instance $x_i \in X$. As a result, we obtain the set of labels $L = \{l_1, \cdots, l_n\}$, with $l_i \in \{1, \cdots, k\}$, that effectively splits $X$ into $k$ non-overlapping clusters $c_i$ to form a partition called $C$. The criterion used to assign an instance to a given cluster is the similarity to the rest of elements in that cluster, and the dissimilarity to the rest of instances of the dataset, which can be obtained with some kind of distance measurement $d(\cdot, \cdot)$. An important detail is that each instance must belong to a single cluster \cite{jain1999data}.

Semi-supervised learning (SSL) is a machine learning paradigm that arises from adding incomplete information to unsupervised learning. We can divide SSL methods into two broad categories according to their objective: semi-supervised classification and semi-supervised clustering. The first method has partial information about the labels, so it tries to minimize the error based on them while taking into account the distribution of unlabeled instances. The second tries to obtain better defined clusters incorporating background information to the clustering process, this is known as constrained clustering and is the main subject of the study presented in this paper \cite{chapelle2009semi, triguero2015self}.

@@

Constraints can be understood in different ways, resulting in three main types of constrained clustering: cluster-level, instance-level and feature-level constrained clustering \cite{bradley2000constrained,davidson2007survey,schmidt2011clustering}. Moreover, hybrid approaches which try to integrate different types of constraints have also been proposed \cite{wang2010clustering}.

@@

Regarding the degree to which the constraints have to be satisfied, we can make a distinction between the concepts of hard and soft constraints. Hard constraints must necessarily be satisfied in the output partition of any algorithm that makes use of them, while soft constraints are taken as a strong guide for the algorithm that uses them but can be partially satisfied in the output partition \cite{seret2014new, wagstaff2001constrained, law2004clustering}. For the purposes of this paper, we will employ the latter.

@@

Differential Evolution (DE) is an evolution-based algorithm that has proven to be excellent in real-domain problem solving \cite{das2011differential}; in this paper we propose the first DE application to find quality solutions for the constrained clustering problem (see Section \ref{sec:SHADE}). In particular, we will take as basis for our proposal one of the DE variants which showed excellent behavior in worldwide competitions, known as SHADE \cite{molina2018insight}. We will make use of the Random-key concept from BRKGA, along with proposing a new fitness function, to build the already mentioned DE variant. We will properly compare the results obtained by our new method with previous nature-inspired algorithms, as well as with the constrained clustering state-of-the-art.

Regarding the organization of this paper, Section \ref{sec:background} reviews the existing knowledge concerning constrained clustering and DE. In Section \ref{sec:brkga} we will briefly review the BRKGA algorithm and its adaptation for constrained clustering. Afterward, the SHADE variant of DE algorithm will be presented in Section \ref{sec:SHADE}. In Section \ref{sec:SHADEadapt} we present the scheme that allow us to apply SHADE to constrained clustering, including the proposal of a new fitness function. Sections from \ref{sec:expSetup} to \ref{sec:analisis} present the experimental setup, results and their analysis respectively. Finally, in Section \ref{sec:conclusiones} we discuss conclusions.

@@

\subsection{Short Introduction to Iterated Local Search}

Before defining the iterated local search---from now on ILS---,it is necessary to understand the local search and the motivation that leads to designing methods with greater exploratory capacity.

A local search is a procedure that consists of performing an heuristics-guided iterative improvement. Taking as initial state a well defined solution to the problem, the local search uses a neighbor generation operator to explore new solutions. When a solution better than the current one is found, it replaces the current one. In order to evaluate the solutions, the local search uses a function that assigns a value to each one of them, this is known as a target function. ***This value will be better the closer the solution that evaluates the optimal solution.***

Both the neighbor generation operator and the target function used by the local search must be defined for each problem. The target function $f(\cdot)$ determines how good the solution achieved is with respect to the purpose of the problem. The neighbor generation operator takes as base a solution to the problem $s$, and generates a new solution $s^\prime$ by modifying some of its components in a local way.

The Algorithm \ref{alg:LS} presents a basic scheme for a local search procedure. It is worth noting that the condition of line 4 must be adapted if the problem to be solved is one of maximization or minimization. In addition, the general scheme considers that the optimization process ends when the current solution is not improved, however the stop condition can be adapted to each problem.

\begin{algorithm}
	\SetNlSty{textbf}{[}{]}
	\SetNlSkip{0.5em}
	\setstretch{1.2}
	\SetKwFunction{GenerateNeighbor}{GenerateNeighbor}
	\SetKwRepeat{Do}{do}{while}
	\KwIn{Initial solution $s$}
	\BlankLine
	\While{$improvement$}{
		$improvement \leftarrow$ \texttt{false} \\

		$s^\prime \leftarrow $ \GenerateNeighbor{$s$}

		\If{$f(s^\prime) < f(s)$}{

			$s \leftarrow s^\prime$\\
			$improvement \leftarrow$ \texttt{true} \\

		}	
	}
	\BlankLine
	\KwRet ($s$)

With the local search defined in this way, it is easy to see that the solution achieved will be the optimal local $s^*$ closest to the initial solution $s$. Figure \ref{img:LS} shows a pictorical representation of the above.

\begin{figure}[!h]
	\centering
	\includegraphics[scale=0.25]{Figures/LS.png}
	\caption{Local optimum achieved by a local search procedure}\label{img:LS}
\end{figure}

It is therefore necessary to appeal to techniques that introduce diversity in the search to avoid the problem of local optimal. This is exactly what ILS is trying to do.

With ILS, once we have reached a local optimum $s^*$, we apply a perturbation to it that, with a high probability, will result in a new solution $s^\prime$ that will not be a local optimum and will be far enough away from $s^*$. After that, we will apply a local search procedure to find a new local optimal $s^{*\prime}$. Figure \ref{img:ILS} shows a representation of the above concepts.

\begin{figure}[!h]
	\centering
	\includegraphics[scale=0.25]{Figures/ILS.png}
	\caption{No se como llamar a esta puta imagen}\label{img:ILS}
\end{figure}

It should be noted that, if the disturbance applied to $s*$ is too small, the most likely is that the local optimum closest to $s^\prime$ is $s^*$ and therefore the solution space will not be explored effectively. On the other hand, if the disturbance is too big, there will be no bias in the generation of new solutions, and therefore the procedure becomes a random reboot-type method.

The method for generating the perturbation applied to the local optimum $s^*$ may vary depending on the problem, and may be based on previous solutions. In this way, it is possible to keep a history of the solutions reached so far, so they can be used when applying the perturbation. Algorithm \ref{alg:ILS} summarizes the ILS overall optimization process.

\begin{algorithm}
	\SetNlSty{textbf}{[}{]}
	\SetNlSkip{0.5em}
	\setstretch{1.2}
	\SetKwFunction{LocalSearch}{LocalSearch}
	\SetKwFunction{Perturbation}{Perturbation}
	\SetKwRepeat{Do}{do}{while}
	\KwIn{Initial solution $s$}
	\BlankLine
	$s^* \leftarrow$ \LocalSearch{$s$}\\
	\While{Termination criteria are not met}{

		$s^\prime \leftarrow $ \Perturbation{$s^*$, $history$}\\
		$s^{*\prime} \leftarrow$ \LocalSearch{$s^\prime$}

		\If{$f(s^{*\prime}) < f(s^*)$}{

			$s^* \leftarrow s^{*\prime}$

		}	
	}
	\BlankLine
	\KwRet ($s*$)

	\caption{Iterated Local Search}\label{alg:ILS}
\end{algorithm}

In the ILS scheme described in Algorithm \ref{alg:ILS} the acceptance criterion only use the difference in the cost of the solutions to be compared to choose between them. This is the most simple scheme and does not depends on previous solutions recorded in the history.

@@

@@

%% CORREGIR DESDE AQUI HASTA EL FINAL

Table \ref{tab:params} shows a summary of the parameter setup used for both BRKGA$_{CC}$ and SHADE$_{CC}$ algorithms.

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.4}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{>{\centering\arraybackslash}c m{5cm} cc}
			\hline
			Parameter & Meaning & BRKGA & SHADE \\
			\hline
			$|P|$ & Population size & 100 & 100 \\
			Evals & Fitness function evaluations & 300000 & 300000 \\
			$P_e$ & Size of the elite set in population & $0.2 * |P|$ & $0.25 * |P|$ \\
			$P_m$ & Number of mutants to be introduced in the population in each generation & $0.2 * |P|$ & - \\
			$p_\text{inherit}$ & Probability that a feature is inherited from an elite parent & $50\%$ & - \\
			$k$ & Output partition number of clusters & \multicolumn{2}{m{4cm}}{No. Classes (Table \ref{tab:datasets}).} \\
			\hline

	\caption{Parameters setup used for BRKGA and SHADE.}
	\label{tab:params}
\end{table}

In both cases---BRKGA$_{CC}$ and SHADE$_{CC}$---the population size will be 100 individuals, and the stop criterion is given by the number of evaluations of the fitness function, which at most will be 300000.

To compare with the state-of-the-art methods mentioned in Section \ref{sec:BackSOTA} we will use the parameters setup shown in Table \ref{tab:paramsSOTA} for the implementation that can be found at \href{https://github.com/GermangUgr/TFG/tree/master/Software}{GitHub}

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.4}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{>{\centering\arraybackslash}c m{5cm} c}
			\hline
			Common Parameters & Meaning & Value \\
			\hline
			\texttt{max\_iter} & Maximum number of iterations & 300 \\
			$k$ & Output partition number of clusters & No. Classes (Table \ref{tab:datasets}). \\
			\hline
			\hline
			Specific Parameters & \multicolumn{2}{l}{Name and Value} \\
			\hline
			COPKM & \multicolumn{2}{l}{\texttt{tolerance} = $1 * 10^{-4}$; \texttt{init\_mode} = \texttt{``rand''}} \\
			CECM & \multicolumn{2}{l}{$\alpha = 1$, $\rho = 100$, $\xi = 0.5$, \texttt{stop\_threshold} = $1 * 10^{-3}$, \texttt{init\_mode} = \texttt{``rand''}} \\
			LCVQE & \multicolumn{2}{l}{\texttt{initial\_centroids} = $\emptyset$} \\
			RDPM & \multicolumn{2}{l}{$\xi_0 = 0.1$, $\xi_\text{rate} = 1$, $\lambda$ is calculated on the basis of the mean distances in the dataset.} \\
			TVClust & \multicolumn{2}{l}{$\alpha_0 = 1.2$, \texttt{stop\_threshold} = $5*10^{-4}$} \\
			\hline

In this section we present Tables from \ref{tab:results10} to \ref{tab:results20SOTA}, which display the results obtained by the methods to be compared for each dataset and constraint set.

Since the methods we are comparing involve non-deterministic procedures, the results may vary from one run to another. To lessen the effect this may have on the results, we will apply each method 5 times to every dataset and constraint set, so that the measures shown in the previously mentioned tables correspond to the average of the 5 runs.

In this section we present a brief comparison between the proposed SHADE$_{CC}$ method and some of the state-of-the-art methods. Tables from \ref{tab:results10SOTA} to \ref{tab:results20SOTA} show the results obtained by these methods. To shorten the presented results this time we will only compare the ARI.

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.4}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{lccccccc}
			\hline
			Dataset & DILS & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
			\hline
			Appendicitis & 0.611 & 0.022 & -1.000 & 0.335 & 0.316 & 0.025 & -0.005 \\
			Breast Cancer & 0.755 & 0.395 & -0.604 & 0.486 & 0.502 & 0.000 & 0.000 \\
			Bupa & 0.870 & 0.281 & -1.000 & -0.005 & -0.005 & -0.004 & -0.011 \\
			Circles & 0.798 & 0.166 & -1.000 & -0.003 & 0.162 & 0.137 & 0.133 \\
			Ecoli & 0.069 & 0.009 & -1.000 & 0.387 & 0.417 & 0.265 & -1.000 \\
			Glass & 0.009 & 0.016 & 0.184 & 0.268 & 0.197 & 0.211 & -1.000 \\
			Haberman & 0.638 & 0.047 & -1.000 & -0.002 & 0.127 & 0.218 & -0.004 \\
			Hayesroth & 0.031 & 0.027 & -1.000 & 0.106 & 0.097 & 0.054 & 0.139 \\
			Heart & 0.793 & 0.483 & -1.000 & 0.027 & 0.036 & 0.423 & -0.003 \\
			Ionosphere & 0.792 & 0.125 & -1.000 & 0.168 & 0.197 & 0.000 & 0.030 \\
			Iris & 0.621 & 0.238 & -0.285 & 0.730 & 0.607 & 0.244 & 0.684 \\
			Led7Digit & 0.016 & 0.009 & 0.497 & 0.425 & 0.369 & 0.316 & -1.000 \\
			Monk2 & 0.815 & 0.421 & 0.982 & 0.072 & 0.094 & -0.002 & 0.007 \\
			Moons & 0.968 & 0.277 & -1.000 & 0.241 & 0.319 & 0.785 & 0.092 \\
			Movement Libras & 0.022 & 0.004 & 0.285 & 0.293 & 0.256 & 0.000 & -1.000 \\
			Newthyroid & 0.028 & 0.017 & -1.000 & 0.579 & 0.289 & 0.846 & 0.002 \\
			Saheart & 0.800 & 0.147 & 0.974 & 0.018 & 0.020 & 0.068 & 0.000 \\
			Sonar & 0.649 & 0.124 & -1.000 & 0.004 & 0.013 & 0.000 & 0.000 \\
			Soybean & 0.304 & 0.496 & 0.503 & 0.545 & 0.621 & 0.000 & 0.244 \\
			Spectfheart & 0.896 & 0.334 & -1.000 & -0.107 & -0.114 & 0.000 & 0.050 \\
			Spiral & 0.856 & 0.058 & -1.000 & -0.003 & 0.012 & 0.034 & -0.002 \\
			Tae & 0.031 & 0.018 & -1.000 & 0.009 & -0.000 & 0.067 & 0.000 \\
			Vowel & 0.002 & 0.003 & -1.000 & 0.063 & -0.003 & 0.067 & -1.000 \\
			Wine & 0.334 & 0.120 & -1.000 & 0.360 & 0.368 & 0.281 & 0.004 \\
			Zoo & 0.151 & 0.105 & 0.715 & 0.666 & 0.412 & 0.335 & -1.000 \\
			\hline

	\caption{Experimental results obtained for $CS_{10}$ comparing SHADE$_{CC}$ and the state-of-the-art.}
	\label{tab:results10SOTA}
\end{table}

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.4}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{lccccccc}
			\hline
			Dataset & DILS & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
			\hline
			Appendicitis & 0.957 & 0.296 & -1.000 & 0.305 & 0.284 & 0.025 & -0.006 \\
			Breast Cancer & 0.788 & 0.772 & 1.000 & 0.486 & 0.502 & 0.000 & 0.037 \\
			Bupa & 0.984 & 0.840 & 1.000 & -0.005 & -0.007 & -0.007 & 0.001 \\
			Circles & 1.000 & 0.828 & 1.000 & -0.003 & 0.375 & 0.973 & 0.105 \\
			Ecoli & 0.134 & 0.015 & -1.000 & 0.387 & 0.372 & 0.719 & -1.000 \\
			Glass & 0.057 & 0.017 & -1.000 & 0.280 & 0.253 & 0.229 & -1.000 \\
			Haberman & 1.000 & 0.866 & 1.000 & -0.002 & 0.075 & 0.219 & -0.050 \\
			Hayesroth & 0.386 & 0.107 & -1.000 & 0.106 & 0.107 & 0.150 & 0.135 \\
			Heart & 1.000 & 0.692 & 1.000 & 0.025 & 0.033 & 0.442 & 0.000 \\
			Ionosphere & 0.970 & 0.803 & 1.000 & 0.178 & 0.212 & 0.000 & 0.057 \\
			Iris & 0.850 & 0.412 & -1.000 & 0.730 & 0.547 & 0.421 & 0.684 \\
			Led7Digit & 0.020 & 0.004 & -1.000 & 0.425 & 0.490 & 0.324 & -1.000 \\
			Monk2 & 0.902 & 0.751 & 1.000 & 0.072 & 0.170 & -0.002 & 0.007 \\
			Moons & 1.000 & 0.958 & 1.000 & 0.241 & 0.436 & 0.987 & 0.095 \\
			Movement Libras & 0.025 & 0.006 & -0.742 & 0.293 & 0.255 & 0.000 & -1.000 \\
			Newthyroid & 0.440 & 0.320 & -1.000 & 0.568 & 0.356 & 0.923 & 0.052 \\
			Saheart & 0.862 & 0.749 & 1.000 & 0.020 & 0.037 & 0.221 & 0.000 \\
			Sonar & 0.981 & 0.800 & -1.000 & 0.004 & 0.019 & 0.000 & 0.000 \\
			Soybean & 0.469 & 0.428 & 0.584 & 0.545 & 0.605 & 0.000 & 0.000 \\
			Spectfheart & 1.000 & 0.924 & 0.983 & -0.107 & -0.117 & 0.000 & -0.070 \\
			Spiral & 1.000 & 0.847 & -1.000 & -0.003 & 0.014 & 0.006 & 0.051 \\
			Tae & 0.299 & 0.119 & -1.000 & 0.008 & -0.004 & 0.062 & 0.000 \\
			Vowel & 0.003 & 0.002 & -1.000 & 0.063 & -0.003 & 0.073 & -1.000 \\
			Wine & 0.725 & 0.528 & -1.000 & 0.360 & 0.367 & 0.348 & -0.001 \\
			Zoo & 0.185 & 0.157 & 0.435 & 0.642 & 0.450 & 0.353 & -1.000 \\
			\hline

		\caption{Experimental results obtained for $CS_{10}$ comparing SHADE$_{CC}$ and the state-of-the-art.}
		\label{tab:results15SOTA}
	\end{table}

>>>>

One of the most widely known and studied data analysis problems is clustering. It is one of the most successful methods in the field of unsupervised learning, where there is no supervision on how the information should be handled. Clustering has been able to provide solutions in a large number of knowledge fields, such as, marketing, banking, psychology, psychiatry, astronomy, archaeology, genetics,\cite{Everitt:2009:CA:1538772} cartographic labeling \cite{araujo2019improving} and image segmentation in various fields, ranging from general applications \cite{wang2018non} to very specific ones such as in medicine \cite{verma2016improved, aparajeeta2016modified}.

In the literature, clustering methods are often divided into two subsets: partitional clustering and hierarchical clustering. In hierarchical clustering the result is not a partition of the data with a certain number of clusters---as in partitional clustering---, but instead a dendrogram in which there are partitions that include from the whole dataset to particular individuals \cite{Everitt:2009:CA:1538772}. A representative example of partitional clustering is the widely studied and well-known algorithm K-means \cite{wu2009top}, while for hierarchical clustering the CURE method should be highlighted \cite{guha1998cure}. In this paper we will focus on partitional clustering.

We can define partitional clustering as the task of grouping the instances of a dataset into $k$ clusters, so that new information can be extracted from them. A dataset $X$ is composed of $n$ instances, each one of them described by $d$ features. More formally, $X = \{x_1, \cdots, x_n\}$, with the $i$th instance noted as $x_i = (x_{[i,1]}, \cdots, x_{[i,u]})$. A typical clustering algorithm assigns a class label $l_i$ to each instance $x_i \in X$. As a result, we obtain the set of labels $L = \{l_1, \cdots, l_n\}$, with $l_i \in \{1, \cdots, k\}$, that effectively splits $X$ into $k$ non-overlapping clusters $c_i$ to form a partition called $C$. The criterion used to assign an instance to a given cluster is the similarity to the rest of elements in that cluster, and the dissimilarity to the rest of instances of the dataset, which can be obtained with some kind of distance measurement $d(\cdot, \cdot)$ \cite{jain1999data}.

Semi-supervised learning (SSL) is a machine learning paradigm that arises from adding incomplete information to unsupervised learning. We can divide SSL methods into two broad categories according to their objective: semi-supervised classification and semi-supervised clustering \cite{chapelle2009semi}. The first method has partial information about the labels, so it tries to minimize the error based on them while taking into account the distribution of unlabeled instances. The second tries to obtain better defined clusters incorporating background information to the clustering process, this is known as constrained clustering and is the main subject of the study presented in this paper \cite{triguero2015self}.

@@

Constraints can be understood in different ways, resulting in three main types of constrained clustering: cluster-level \cite{bradley2000constrained}, instance-level \cite{davidson2007survey} and feature-level constrained clustering \cite{schmidt2011clustering}. Moreover, hybrid approaches which try to integrate different types of constraints have also been proposed \cite{wang2010clustering}.

@@

Regarding the degree to which the constraints have to be satisfied, we can make a distinction between the concepts of hard \cite{wagstaff2001constrained} and soft \cite{law2004clustering} constraints. Hard constraints must necessarily be satisfied in the output partition of any algorithm that makes use of them, while soft constraints are taken as a strong guide for the algorithm that uses them but can be partially satisfied in the output partition \cite{seret2014new}. For the purposes of this paper, we will employ the latter.

@@

Differential Evolution (DE) is an evolution-based algorithm that has proven to be excellent in real-domain problem solving \cite{das2011differential}. In this paper we propose the first DE application to find quality solutions for the constrained clustering problem. In particular, we will take as basis for our proposal one of the DE variants which showed excellent behavior in worldwide competitions, known as SHADE \cite{molina2018insight}. We will make use of the Random-key concept from BRKGA, along with proposing a new fitness function, to build the already mentioned DE variant. We will properly compare the results obtained by our new method with previous nature-inspired algorithms, as well as with the constrained clustering state-of-the-art.

Regarding the organization of this paper, Section \ref{sec:background} reviews the existing knowledge concerning constrained clustering and DE. In Section \ref{sec:brkga} the BRKGA algorithm and its adaptation for constrained clustering will be briefly reviewed. Afterward, the SHADE variant of DE algorithm will be presented in Section \ref{sec:SHADE}. In Section \ref{sec:SHADEadapt} the scheme that allow us to apply SHADE to constrained clustering, including the proposal of a new fitness function, will be presented. Sections from \ref{sec:expSetup} to \ref{sec:analisis} present the experimental setup, results and their analysis respectively. Finally, in Section \ref{sec:conclusiones} conclusions are discussed.

\textcolor{red}{Decir que mientras que en los algoritmos memeticos el host es el GA en DILS el host es ILS, conservando todas sus caracteristicas y añadiendo algunas tecnicas de los GA}

@@

\textcolor{red}{Poner aqui una descripción del BRKGA?}

@@

The Dual Iterative Local Search (DILS) is a new variant of the classic Iterated Local Search (ILS) method. Its goal is to perform a search in the solutions space to find the solution with the best fitness function $f$ value by introducing diversity in a guided way. While ILS works with a single individual, DILS keeps two of them $\{m_b, m_w\}$ in memory at all times, which allows it to guide diversity-introducing methods and to avoid local optima. The individual $m_b$ is the one providing the best fitness value, so $m_w$ provides the worst fitness value.

DILS mimics some of the genetic algorithms techniques to introduce diversity into the optimization process, while also manteining the esentials of ILS. In each generation $G$ of the optimization process DILS builds a new individual by applying a crossover operator to $m_b$ and $m_w$. After that, it applies a strong mutation operator to the newly generated individual. This is the invidual DILS applies the local search procedure to in order to improve its fitness value, resulting in the trail individual $m_t$. The trial individual $m_t$ replace the worst of its parents $m_w$ if it provides a better fitness value, this operation makes the replacement operator for DILS. Equation \ref{eq1} displays the expression for this operator in the case of a minimization problem.

\begin{equation}
m_{w,G} = \left\{ \begin{array}{lc}
m_t &   \text{if} \;\; f(m_t) < f(m_w)\\
m_{w,G} &  \text{otherwise}
\end{array}
\right..
\label{eq1}
\end{equation}

Now, the process described so far will most likely fall into a local optimum, effectively stopping the exploration of the solutions space. To avoid it DILS implements a reinitialization method for $m_w$ based on the difference between the two individuals it keeps in memory $\{m_b, m_w\}$, Equation \ref{eq2} shows the reinitialization criterion for a minimization problem.

\begin{equation}
m_{w,G+1} = \left\{ \begin{array}{lc}
\text{RandInit}() &   \text{if} \;\; f(m_b) - f(m_w) > f(m_b) * \xi \;\;\\
m_{w,G} &  \text{otherwise}
\end{array}
\right.,
\label{eq2}
\end{equation}

\noindent where RandInit$()$ is a function that returns a randomly initialized individual and $\xi \in [-1,1]$ is the parameter that controls the tolerance of the reinitialization method. This way the best individual is always preserved and therefore always takes part in the process of generating new individuals (via crossover operator) as the best parent. Figure \ref{img:DILS_diag} summarizes the DILS optimization process.

\begin{figure}[!h]
	\centering
	\includegraphics[scale=0.25]{Figures/DILSDiagram.png}
	\caption{Diagram summarizing DILS optimization process. \textcolor{red}{La figura es horrible (llamar a javi)}}\label{img:DILS_diag}
\end{figure}

It is crucial to note that in Figure \ref{img:DILS_diag} arrows in red indicate stages of the algorithm where $m_b$ and $m_w$ are reasigned to the true best and worst individual respectively. This way, at the point where the reinitialization criterion is tested, it is possible that $f(m_w)$ provides a better value than $f(m_b)$, so that $f(m_b) - f(m_w)$ can be either positive or negative.

This leads us to note the influence of the parameter $\xi$. In one hand, if we are dealing with a minimization problem and we set $\xi > 0$, then the true worst individual is only reinitializatized when the trial individual $m_t$ replaced $m_w$ and is also better than $m_b$ by a margin. On the other hand and also for the minimization case but this time with $\xi < 0$ we allow DILS to reinitialize the true worst individual even $m_t$ is not better than $m_b$ but they are close. It comes clear that when $\xi = 0$ the worst individual is restarted if $f(m_b) = f(m_w)$. Note that $\xi$ could take any real value, but it is not realistic to set it outside of the range $[-1,1]$. To summarize, the closer $\xi$ is to 1 the more restrictive the reinitialization criterion is---in the minimization case. It is also worth noting that in early stages of the optimization process the reinitialization criterion tends to be met more often than in later stages, resembling some fatures of the simmulated annealing algorithm \cite{van1987simulated}. Algorithm \ref{alg:DILS} summarizes the overall DILS optimization process.

\begin{algorithm}
	\SetNlSty{textbf}{[}{]}
	\SetNlSkip{0.5em}
	\setstretch{1.2}
	\SetKwFunction{RandInit}{RandInit}
	\SetKwFunction{Best}{Best}
	\SetKwFunction{Worst}{Worst}
	\SetKwFunction{Crossover}{Crossover}
	\SetKwFunction{Mutation}{Mutation}
	\SetKwFunction{LocalSearch}{LocalSearch}
	\KwIn{}
	\tcp{Initialization phase}
	$G \leftarrow 0$\\
	$m_b \leftarrow$ \RandInit{};
	$m_w \leftarrow$ \RandInit{}\\
	\tcp{Main loop}
	\While{Termination criteria are not met}{
		\tcp{Find best and worst individual}
		$m_b \leftarrow $ \Best{$ \{m_b, m_w\} $};
		$m_w \leftarrow $ \Worst{$ \{m_b, m_w\} $}\\
		\tcp{Generate new individual}
		$m_t \leftarrow $ \Crossover{$m_b, \; m_w$}\\
		$m_t \leftarrow $ \Mutation{$m_t$}\\
		\tcp{Improve new individual}
		$m_t \leftarrow $ \LocalSearch{$m_t$}\\
		\tcp{Apply replacement operator}
		\If{$f(m_t) < f(m_w)$}{

			$m_w \leftarrow m_t$\\

		}
		\tcp{Check restart criterion}
		\If{$f(m_b) - f(m_w) > f(m_b) \times \xi$}{

			$m_b \leftarrow $ \Best{$ \{m_b, m_w\} $}\\
			$m_w \leftarrow$ \RandInit{}\\

		}

	$G \leftarrow G + 1$
}
\caption{DILS}\label{alg:DILS}
\end{algorithm}

For the mutation and crossover operators that appear in the Algorithm \ref{alg:DILS}, we use the uniform crossover operator \cite{spears1995virtues} and the segment mutation operator \textcolor{red}{REFERENCIA?} respectively. This type of mutation consists of selecting a fixed size segment from the features of the individual and replacing it with a random segment of the same size. The size of the segment must be fixed at the beginning of the optimization process.

Cabe destacar que el proceso de optimización descrito hasta el momento es aplicable a cualquier problema de minimización, aunque es claro que axiste una adaptación directa de DILS para problemas de maximización. Además el proceso de búsqueda local empleado para obtener el individuo trial debe ser especificado para cada problema, así como cuestiones de representación numérica y las criterios de parada.

It should be noted that the optimization process described so far is applicable to any minimization problem, although it is clear that there is a direct adaptation of DILS for maximization problems. In addition, the local search procedure used to obtain the trial individual $m_t$ must be specified for each problem, as well as numerical representation matters and termination criteria.

\section{DILS optimization scheme for constrained clustering}

In this section, we present an optimization scheme for the DILS algorithm when applied to constrained clustering. We will call this new approach DILS$_{CC}$. To begin with, we will use an integer-based representation for all the individuals DILS$_{CC}$ work with; this way, each individual $m$ is defined as a vector of $n$ integers---being $n$ the number of instances of the dataset. Each integer $m_i \in [0,k)$ in $m$ represent the label of the cluster that instance $x_i$ belongs to, si each individual $m$ is a solution to the clustering problem. Additionally, in order to not bias the exploration of the solution space, we will use a random initialization to intially set $m_b$ and $m_w$, this procedure will also be used when $m_w$ must be reinitialized.

Regarding the fitness function $f$, for each individual $m$ its fitness value $f_m$ can be calculated as in Equation \ref{eq3}.

\begin{equation}
f_m = z_m * (\text{infeasibility}_m + 1),
\label{eq3}
\end{equation}

\noindent where $\text{infeasibility}_m$ is the number of non-satisfied constraints and $z_m$ is the within-cluster-sum-of-squares which can be computed as in Equation \eqref{eq4}.

\begin{equation}
z_m = \sum_{c_i \in C_m} \left[ \frac{\sum_{x_a, x_b \in c_i} d^2(x_a,x_b)}{|c_i|}\right].
\label{eq4}
\end{equation}

We also need to specify a local search procedure for DILS to be able to improve the mutant individual to obtain the trial vector $m_t$. Algorithm \ref{alg:LS} shows the local search precedure used in DILS$_{CC}$. Note that, in order to not to bias the search, indices from vector $m$ and labels from the set of labels are randomly chosen to generate a random subset of the neighborhood of $m$. We also stablish a maximum number of neighbors $max\_neighbors$ that can be generated, this is done to find a better exploration-exploitation balance.

\begin{algorithm}
	\SetNlSty{textbf}{[}{]}
	\SetNlSkip{0.5em}
	\setstretch{1.2}
	\SetKwFunction{RandomShuffle}{RandomShuffle}
	\SetKwFunction{Rand}{Rand}
	\SetKwRepeat{Do}{do}{while}
	\KwIn{Dataset $X$, constraint sets $C_=$ and $C_{\neq}$, individual (solution) $m$, number of clusters $k$, maximum number of neighbors that can be generated $max\_neighbors$.}
	\BlankLine
	\While{$improvement$ {\bf and} $generated < max\_neighbors$}{
		$improvement \leftarrow$ \texttt{false} \\
		\tcp{Select random index (object) from $m$}
		$i \leftarrow $ \Rand{$1$,$n$} \\
		\tcp{Random shuffle labels set}
		$RSL \leftarrow $ \RandomShuffle{$\{1,\cdots,K\}$}\\
		\For{$l \in RSL$}{
			$m^\prime \leftarrow m$\\
			\tcc{Move object $i$ from $m$ to the cluster associated with label $l$}
			$m^\prime_i \leftarrow l$\\
			\If{$f(m^\prime) < f(m)$}{
				$m \leftarrow m^\prime$\\
				$improvement \leftarrow$ \texttt{true} \\
			}
			$generated \leftarrow generated + 1$\\
		}
	}
	\BlankLine
	\KwRet ($m$)

@@

\subsection{Validation of results} \label{sec:ValidtnMethod}

In order to validate the results which will be presented in Section \ref{sec:results} we will use Bayesian statistical tests, instead of the classic Null Hypothesis Statistical Tests (NHST). In \cite{benavoli2017time} we find an in-depth analysis of the disadvantages of NHST, and a new model is proposed for carrying out comparisons researchers are interested in. \textit{"In a nutshell: NHST do not answer the question we ask"}. To put it clear, the disadvantages of the NHST that the authors highlight in \cite{benavoli2017time} are based on the trap of black-and-white thinking, this is: to reject, or not to reject?
